---
title: "Cross-lag SEM models with Lavaan"
author: "Michelle.VanTieghem"
date: "Jan 23, 2019"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
---


## get useful packages
```{r, include = F, warnings = F, echo = F}
library(lavaan)
library(tidyverse)
```

## function for chi-square test later
```{r}
 # is p < 0.05? test with the chi-square distribution for this DF
chi_square_test <- function(chi_sq_diff, df_diff) {
  if (chi_sq_diff > qchisq(.95, df=df_diff)){
    p  = '< 0.05'
  } else{
    p = '> 0.05' 
  }
  return(list(chi_square_difference = chi_sq_diff, df_difference = df_diff, p.value = p))
}

```

## load data 
```{r}

# this one compiled more subjects with 2 waves of data. 
load("../../../data/5_SEM_cross_lag_data/cleaned2_all_2wave_cort_brain_wide_format_ages4to19.Rdata")
nrow(df_wide)
```

# check for correlations between amyg and hipp
```{r}

amyg_hipp_cor1 <- cor.test(df_wide$Amyg_wave1, df_wide$Hipp_wave1)
save(amyg_hipp_cor1, file = "model_output/amyg_hipp_correlation_wave1.Rdata")
amyg_hipp_cor2 <- cor.test(df_wide$Amyg_wave2, df_wide$Hipp_wave2)
save(amyg_hipp_cor2, file = "model_output/amyg_hipp_correlation_wave2.Rdata")

```


# cross-lagged SEM Cort & Hipp 
## basic model with cross-lag paths 

note on standardized coefficients
Two extra columns of standardized parameter values are printed. In the first column (labeled Std.lv), only the latent variables are standardized. In the second column (labeled Std.all), both latent and observed variables are standardized. The latter is often called the 'completely standardized solution'.
```{r }

model= 
'
#Regression paths cross-lagged
Hipp_wave2_rs.c ~  Hipp_wave1_rs.c + waking_wave1.c + Age_wave2.c +  IDENT_SUBTYPE + sex.c + ICV_wave2_rs.c
waking_wave2.c ~ Hipp_wave1_rs.c + waking_wave1.c + Age_wave2.c + IDENT_SUBTYPE + sex.c 

# Regression paths at T1 
Hipp_wave1_rs.c ~ Age_wave1.c + IDENT_SUBTYPE  +  sex.c + ICV_wave1_rs.c 
waking_wave1.c ~ Age_wave1.c + IDENT_SUBTYPE +  sex.c 

# covariance
waking_wave1.c ~~ Hipp_wave1_rs.c
waking_wave2.c ~~ Hipp_wave2_rs.c

# time 1 - time 2 things
ICV_wave1_rs.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Age_wave2.c 

# correlations at each time point 
Age_wave1.c ~ ICV_wave1_rs.c 
Age_wave2.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Hipp_wave1_rs.c 
Age_wave2.c ~ Hipp_wave2_rs.c
Age_wave1.c ~ waking_wave1.c
Age_wave2.c ~ waking_wave2.c
#sex.c	~	Hipp_wave1_rs.c		

ICV_wave1_rs.c	~	Hipp_wave1_rs.c	
ICV_wave2_rs.c	~	Hipp_wave2_rs.c		

IDENT_SUBTYPE	~	Hipp_wave1_rs.c
#IDENT_SUBTYPE ~ sex.c 
IDENT_SUBTYPE ~ ICV_wave1_rs.c
'

hipp_mod1 <- sem(model, data=df_wide, fixed.x=F, missing='fiml')

fit_measures <- data.frame(fitMeasures(hipp_mod1)) 
names(fit_measures) <- "value"
fit_measures$metric <- rownames(fit_measures) 
fit_measures$model <- "hipp_mod1"

hipp_mod1_fit_measures <- fit_measures %>%
  spread(metric, value)

summary(hipp_mod1, fit.measures=TRUE, standardized=TRUE, rsquare=TRUE)
# parameterEstimates(hipp_mod1) just useful for confidence intervals.

```

## testing model without hipp1 --> waking 2
```{r }

model= 
'
#Regression paths cross-lagged
Hipp_wave2_rs.c ~  Hipp_wave1_rs.c + waking_wave1.c + Age_wave2.c +  IDENT_SUBTYPE + sex.c + ICV_wave2_rs.c 
waking_wave2.c ~  waking_wave1.c + Age_wave2.c + IDENT_SUBTYPE + sex.c 

# Regression paths at T1 
Hipp_wave1_rs.c ~ Age_wave1.c + IDENT_SUBTYPE  +  sex.c + ICV_wave1_rs.c 
waking_wave1.c ~ Age_wave1.c + IDENT_SUBTYPE +  sex.c 

# covariance
waking_wave1.c ~~ Hipp_wave1_rs.c
waking_wave2.c ~~ Hipp_wave2_rs.c

# time 1 - time 2 things
ICV_wave1_rs.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Age_wave2.c 

# correlations at each time point 
Age_wave1.c ~ ICV_wave1_rs.c 
Age_wave2.c ~ ICV_wave2_rs.c 
Age_wave1.c ~ Hipp_wave1_rs.c 
Age_wave2.c ~ Hipp_wave2_rs.c
Age_wave1.c ~ waking_wave1.c
Age_wave2.c ~ waking_wave2.c
#sex.c	~	Hipp_wave1_rs.c		

ICV_wave1_rs.c	~	Hipp_wave1_rs.c	
ICV_wave2_rs.c	~	Hipp_wave2_rs.c		

IDENT_SUBTYPE	~	Hipp_wave1_rs.c
#IDENT_SUBTYPE ~ sex.c 
IDENT_SUBTYPE ~ ICV_wave1_rs.c
'

hipp_mod2 <- sem(model, data=df_wide, fixed.x=F, missing='fiml')

fit_measures <- data.frame(fitMeasures(hipp_mod2)) 
names(fit_measures) <- "value"
fit_measures$metric <- rownames(fit_measures) 
fit_measures$model <- "hipp_mod2"

hipp_mod2_fit_measures <- fit_measures %>%
  spread(metric, value)
```

## chisquare test of model fit 
adding hipp wave1 --> cort wave 2 significantly improves model fit

```{r}
hipp_mod1_fit_measures$chisq # model with hipp has better (lower) chi-square
hipp_mod2_fit_measures$chisq # model without hipp has higher (worse) chi-square 

# is this difference significant?
diff_chi_sq <- abs(hipp_mod2_fit_measures$chisq - hipp_mod1_fit_measures$chisq)
diff_df <-  abs(hipp_mod2_fit_measures$df - hipp_mod1_fit_measures$df)

# yes!
chi_square_hipp_result <- chi_square_test(diff_chi_sq, diff_df)
chi_square_hipp_result
save(chi_square_hipp_result, file = "hipp_cross_lag_chi_sq_test.Rdata")

```

# save final model 

```{r}

save(hipp_mod1, file = "model_output/hipp_waking_cross_lagged_model_main_effects.Rdata")

```


