---
title: "making a stacked df for amyg-cort with covariates "
author: "Michelle.VanTieghem"
date: "June 12, 2018"
output: 
    html_document:
        toc: yes
        toc_float:
            smooth_scroll: true
            collapsed: false
        number_sections: false
--- 

```{r, warnings = F, include = F}

library(tidyverse)
library(readxl)
```

# load data file for longitudinal cort. 
updated file from June 2018 corrections from Marta and added testosterone. 
```{r, warnings = F}
load( "../../..data/1_NEW_cort_and_T_marta_version_data/Marta_v_2A_long_cort_testos_cleaned.Rdata")
```

## make a SUBJECTID_long variable to use for merging later
```{r}

cort_merge8$SUBJECTID_long <- ifelse(cort_merge8$index_wave == 2, 
                                      paste0(cort_merge8$IDENT_SUBID, "_fu1"),
                               ifelse(cort_merge8$index_wave == 3, 
                                      paste0(cort_merge8$IDENT_SUBID, "_fu2"), 
                                      as.character(cort_merge8$IDENT_SUBID)))

```


## clean this up a little so that only essential variables are included. 
```{r}
cort_df <- cort_merge8 %>%
  select(
    # demographic variables that will match brain stuff. 
    IDENT_SUBID, SUBJECTID_long, IDENT_SUBTYPE, DEM_3_GENDER_CHILD,  corrected_cort_age_yrs, 
    # actual variable of interest (after excluding outliers.)    
     cort_values_fixed_no_outliers,
         # timing variables 
         index20, index4, time_of_day,index_day, time_of_day, index_wave, batch,
    # covariates that are impt to keep.
        testos_cleaned, meds_by_wave, PDS_mean, person.centered.time)
```


# add freesurfer data
note: this is already cleaned, and ages 4-19. 

## loading 
```{r}
load("../../../data/2_fs_data/3_cleaned_fsdata_for_growth_chart_brain_age_yrs4_19.Rdata")
#names(fsdata6)
```

## clean up like other dataset
```{r}

fsdata6$IDENT_SUBTYPE <- ifelse(fsdata6$GROUP == "PI", 1, 0)
brain_df <- fsdata6 %>% 
  select(
    # demographics 
    IDENT_SUBID, SUBJECTID_long, IDENT_SUBTYPE, DEM_3_GENDER_CHILD, brain_age_yrs, 
        # variable of interest 
    Hipp_ave, Amyg_ave, 
    # timing variable
    index_wave,
    # covariates 
    ICV, motion_ave_new, aseg_ave_rating)


# check that no missing data. 
sum(is.na(brain_df$ICV))
nrow(brain_df) # should be 288 columns with 

# RESCALE 
brain_df$Amyg_rescaled <- brain_df$Amyg_ave / 1000

```

## stack cort and brain data 

```{r}
# take only essential variables here for now. 
cort_merge <- cort_df %>%
  filter(time_of_day == "waking") %>%
  select(IDENT_SUBID, SUBJECTID_long, index_wave, cort_values_fixed_no_outliers) 

# take everything else as covariates, BUT NOT THE NESTED THINGS (index4, index20, time of day, etc)
cort_covariates <- cort_df %>% 
  filter(time_of_day == "waking") %>%
  select(IDENT_SUBID, SUBJECTID_long, index_wave, IDENT_SUBTYPE, batch, PDS_mean, testos_cleaned, 
         meds_by_wave, corrected_cort_age_yrs) %>% 
  group_by(IDENT_SUBID, index_wave) %>%
  summarize( #index_wave = mean(as.numeric(as.character(index_wave))), 
             IDENT_SUBTYPE = mean(IDENT_SUBTYPE),
             batch = mean(batch, na.rm = T),
             PDS_mean = mean(PDS_mean, na.rm = T),
             testos_cleaned = mean(testos_cleaned, na.rm = T), 
             meds_by_wave = mean(as.numeric(as.character(meds_by_wave)), na.rm = T),
             corrected_cort_age_yrs = mean(corrected_cort_age_yrs), na.rm = T)


# now only take brain values 
brain_merge <- brain_df %>% 
  select(IDENT_SUBID, SUBJECTID_long, index_wave, Amyg_rescaled)


# save the covariates 
brain_covariates <- brain_df %>% 
  select( -Amyg_rescaled, -DEM_3_GENDER_CHILD)


# rename Amyg_rescaled and Hipp_rescaled so they are both labeled 'outcome'
# so that they have identical column names for stacking names(brain_merge)[4] <- "outcome"
names(cort_merge)[4] <- "outcome"
names(brain_merge)[4] <- "outcome"

# rename this so clear what it indicates 
brain_merge$cort_indicator <- 0
cort_merge$cort_indicator <- 1

# use RBIND to stack the variables on top of eachother 
# this will only work if they have identical column names 
stack_df <- rbind(cort_merge, brain_merge)

# now add covariates in 
# when adding cort covariates, values will be NA for brain rows
stack_df3 <- merge(stack_df2, cort_covariates, by = c("IDENT_SUBID", "index_wave"), all = T)

# when adding brain covariates, values will be NA for cort rows. 
stack_df4 <- merge(stack_df3, brain_covariates, by =  c("IDENT_SUBID", "index_wave"), all = T)

```

## modifying covariates to change NA to zeros. 
```{r}
# brain only 
stack_df4$ICV[is.na(stack_df4$ICV)] <- 0
stack_df4$aseg_ave_rating[is.na(stack_df4$aseg_ave_rating)] <- 0
stack_df4$motion_ave_new[is.na(stack_df4$motion_ave_new)] <- 0
stack_df4$scanner_confound <- ifelse(stack_df4$index_wave == 3, 1, 0)
sum(is.na(stack_df4$scanner_confound))

# cort only 
stack_df4$meds_by_wave[is.na(stack_df4$meds_by_wave)] <- 0
stack_df4$index20[is.na(stack_df4$index20)] <- 0
stack_df4$batch[is.na(stack_df4$batch)] <- 0
stack_df4$index_day[is.na(stack_df4$index_day)] <- 0

```

## add more variables you need from master! 
only getting partial master data, stuff i will need.
```{r}

master <- read.csv("../../../data/master/J.Data_Entry_Master_8_17_2017_fixed_fixed.csv")

master2 <- master %>% 
  select(IDENT_SUBID, IDENT_SUBTYPE, DEM_3_GENDER_CHILD)

# add in these variables into long format! 
# note not using "full" option here, because don't want anyone included 
# in the dataset who doesn't have at least ONE VARIABLE of interest.
merged_brms <- merge(stack_df4, master2, by = "IDENT_SUBID")

```

## save data 
this has one row per wave per subject
```{r}

save(merged_brms, file="../../../data/8_mixed_effects_brms/cort_super_long_Amyg_long_stacked_df_2019-01-19.Rdata")

```

