---
title: "mixed effects SEM style with BRMS"
author: "Michelle.VanTieghem"
date: "10/9/2018"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

# Read me 
attempting to test niall's multilevel SEM through random effects in BRMS 
cores = 4 argument to run parallel!
updating Jan 12, 2018 to change model specs to separate between and within subj age.


# load data 
```{r}

load(file="../../../data/8_mixed_effects_brms/cort_super_long_Amyg_long_stacked_df_2019-01-19.Rdata")
head(merged_brms)

library(tidyverse)
library(brms)
```


# make age variables 
## age per wave (across cort and brain because may have one or the other)
```{r}
merged_brms$AGE <- with(merged_brms, # if we have scan age, use that, else, use cort age.
                        ifelse(!is.na(brain_age_yrs), brain_age_yrs, corrected_cort_age_yrs))
sum(is.na(merged_brms$AGE))

merged_brms$AGE.c <- merged_brms$AGE - mean(merged_brms$AGE)
hist(merged_brms$AGE.c)

```

### age at baseline (age intercept)
we do want to mean center this, so average age = 9.8
```{r}
sublist <- unique(merged_brms$IDENT_SUBID)
length(sublist)
merged_brms$Age_wave1 <- NA
merged_brms$index_wave <- as.numeric(as.character(merged_brms$index_wave))

for (sub in sublist){
#  print(sub)
  min_wave <- merged_brms %>%
    group_by(IDENT_SUBID) %>%
    filter(IDENT_SUBID == sub)  %>%
    summarize(min_wave = min(index_wave))
   keep_age <- merged_brms$AGE[merged_brms$index_wave == min_wave$min_wave & merged_brms$IDENT_SUBID == sub]
  merged_brms$Age_wave1[merged_brms$IDENT_SUBID == sub] <- keep_age[1]
}

# mean center at age 9.8
mean(merged_brms$Age_wave1)
merged_brms$Age_wave1.c <- merged_brms$Age_wave1 - mean(merged_brms$Age_wave1)
hist(merged_brms$Age_wave1)
```

### age change (within subject)
 we do not want to mean center this, we want 0 = baseline.
```{r}
# for whatever row is whatever wave, subtract it from the baseline age for that subject 
# to get how many years later that wave of data happened (or 0)
merged_brms$Age_change <- merged_brms$AGE - merged_brms$Age_wave1
mean(merged_brms$Age_change)

# check that this worked 
wave1 <- merged_brms %>%
  filter(index_wave == 1) 

summary(wave1$Age_change)

# wave 2 may be baseline for some subjects, so min =0 
wave2 <- merged_brms %>%
  filter(index_wave == 2) 

summary(wave2$Age_change)

# wave 3 may be baseline for some subjets.
wave3 <- merged_brms %>%
  filter(index_wave == 3) 

summary(wave3$Age_change)
```


## need to make indicator variables 
```{r}
# indicator for each set of data 
merged_brms$CORT <- with(merged_brms, ifelse(indicator_cort ==1, 1, 0))
merged_brms$BRAIN <- with(merged_brms, ifelse(indicator_cort == 0, 1, 0))

# check indicator variables
summary(as.factor(merged_brms$CORT))
summary(as.factor(merged_brms$BRAIN))

# make indicator X Age intercept variables 
merged_brms$CORTxAge_wave1.c <- merged_brms$CORT*merged_brms$Age_wave1.c
merged_brms$BRAINxAge_wave1.c <- merged_brms$BRAIN*merged_brms$Age_wave1.c

# make indicator X Age change variables 
merged_brms$CORTxAge_change<- merged_brms$CORT*merged_brms$Age_change
merged_brms$BRAINxAge_change <- merged_brms$BRAIN*merged_brms$Age_change

```


## center all covariates and check no NAs
even for covariates only relevant to cort, should still be ZERO (ie. batch)
even for covariates only relevant to brain, should still be zero (i.e. ICV)
```{r}

merged_brms$ICV <- merged_brms$ICV/ 10000 # rescale first.
merged_brms$ICV.c <- merged_brms$ICV - mean(merged_brms$ICV, na.rm = T)

merged_brms$sex.c <- merged_brms$DEM_3_GENDER_CHILD - mean(merged_brms$DEM_3_GENDER_CHILD)
sum(is.na(merged_brms$sex.c))

# mean center batch covariate. 
merged_brms$batch.c <- as.numeric(as.character(merged_brms$batch)) - mean(as.numeric(as.character(merged_brms$batch)), na.rm = T)

sum(is.na(merged_brms$batch))
sum(is.na(merged_brms$ICV))
```


## model with cort + brain
no correlation of random effects. 
```{r}

model1 <- brm(outcome ~ 
                # NEED the -1 to remove the intercept 
      # so it will calculate separate intercepts for age and cort and reduce the colinearity!! 
                -1 + 
              # intercept, fixed effects of age, fixed effects of age change
              CORT + CORTxAge_wave1.c +  CORTxAge_change +
              BRAIN + BRAINxAge_wave1.c + BRAINxAge_change +
              # covariates
                sex.c +
              # random intercepts and slopes for age change within subject
                (-1 +  CORT + CORTxAge_change +
                BRAIN +  BRAINxAge_change | IDENT_SUBID), 
              data = merged_brms, chains = 4, cores = 4, 
             control = list(max_treedepth = 15))

#save(model1, file = "brms_cort_brain_stacked_with_age_int_and_age_change.Rdata")
#load("~/Documents/Columbia/LabDATA/Projects/git_repos/structural_analysis_2018/scripts/not_manuscript/8_mixed_effects_brms/amygdala/models/brms_cort_brain_stacked_with_age_int_and_age_change.Rdata")

summary(model1)$random

```

## Add covariates to  model 
```{r}
model2 <- brm(outcome ~ 
                # NEED the -1 to remove the intercept 
            # so it will calculate separate intercepts for age and cort and reduce the colinearity!! 
                -1 + 
              # intercept, fixed effects of age, fixed effects of age change
              CORT + CORTxAge_wave1.c +  CORTxAge_change +
               BRAIN + BRAINxAge_wave1.c + BRAINxAge_change +
               # variable-dependent covariates
                BRAIN*ICV.c +  BRAIN*scanner_confound + 
              CORT*batch.c + CORT*meds_by_wave +
                # covariate for all data
               sex.c +
               (-1 +  CORT + CORTxAge_change +
               BRAIN +  BRAINxAge_change | IDENT_SUBID), 
              data = merged_brms, chains = 4, cores = 4, 
             control = list(max_treedepth = 15))


save(model2, file = "brms_cort_brain_stacked_with_age_int_and_age_change_with_covariates.Rdata")
#load(file = "models/brms_cort_brain_stacked_with_age_int_and_age_change_with_covariates.Rdata")
summary(model2)
```


