---
title: "mixed effects SEM style with BRMS"
author: "Michelle.VanTieghem"
date: "10/9/2018"
output: html_document
---

# Readme 
attempting to test niall's multilevel SEM through random effects in BRMS 
use cores = 4 argument to run in parallel!! 

useful info:
# pulling out fixed effects only 
fixef(test) 
# pulling out random effects
ranef(test)
## pulling out both combined 
coef(test)

# load data

```{r}

load(file="../../../../data/archived/8_mixed_effects_brms/super_long_Hipp_Amyg_long_stacked_df.Rdata")
# check data structure 
head(merged_brms)
```

```{r}
# load packages 
library(lme4)
library(tidyverse)
library(brms) # for baysian models
library(shinystan) # not sure if u need this?
```

## clean data more
```{R}
## need to make a scanner coefficient to control for 
merged_brms$scanner_confound <- as.factor(ifelse(merged_brms$index_wave == 3, 1, 0))
summary(merged_brms$scanner_confound)

##  make a centered- age variable 
merged_brms$AGE <- merged_brms$HIPP_age_yrs.x

# mean age = 10.5
merged_brms$AGE.c <- merged_brms$AGE - mean(merged_brms$AGE)
hist(merged_brms$AGE.c)

## make indicator variables 
merged_brms$AMYG<- with(merged_brms, ifelse(indicator_amyg ==1, 1, 0))
merged_brms$HIPP <- with(merged_brms, ifelse(indicator_amyg == 0, 1, 0))

# check indicator variables
summary(as.factor(merged_brms$AMYG))
summary(as.factor(merged_brms$HIPP))

# make indicator X Age variables 
merged_brms$AMYGxAGE.c <- merged_brms$AMYG*merged_brms$AGE.c
merged_brms$HIPPxAGE.c <- merged_brms$HIPP*merged_brms$AGE.c

## center all covariates 
merged_waking$ICV <- merged_waking$ICV.x/ 10000 # rescale first.
merged_waking$ICV.c <- merged_waking$ICV.x - mean(merged_waking$ICV.x, na.rm = T)
merged_waking$testos_cleaned.c <- merged_waking$testos_cleaned - mean(merged_waking$testos_cleaned, na.rm = T)
merged_waking$sex.c <- merged_waking$DEM_3_GENDER_CHILD.x - mean(merged_waking$DEM_3_GENDER_CHILD.x)
```

## model with AMYG + HIPP
once i included the fixed effects properly, there is no correlation of random effects. 
```{r}

model1 <- brm(outcome ~ 
                       # note: -1 is necessary to remove the overall model intercept 
                    # so it will calculate separate intercepts for age and AMYG and reduce the colinearity!! 
                       -1 + 
                       # fixed effects for amygdala intercept/change, hipp intercept/change and covariates.
                       AMYG + AMYGxAGE.c + HIPP + HIPPxAGE.c + sex.c + IDENT_SUBTYPE + 
                       
                       # random effects for amygdala intercept/change and hipp intercept/change 
                   (-1 + AMYG + AMYGxAGE.c + HIPP + HIPPxAGE.c | IDENT_SUBID), 
                    data = merged_waking, 
                   # recommended settings for chains = 4
                   # cores = 4 means it will run on 4 cores on your computer (if possible)
                   chains = 4, cores = 4, 
                   # control options will help if model doesn't converge.
                    control = list(max_treedepth = 15))
# save 
save(model1, file = "models/model1_results.Rdata")
summary(model1) 
```


## pull out the model correlations for random effects
showing whether individual intercept/slopes between amyg and hipp are correlated.
```{r}
mod_ranef_corr <- summary(model1)$random
mod_cor_AMYG_AMYGxAGE.c <- mod_ranef_corr$IDENT_SUBID[5]
mod_cor_AMYG_HIPP <- mod_ranef_corr$IDENT_SUBID[6]
mod_cor_AMYGxAGE.c_HIPP <- mod_ranef_corr$IDENT_SUBID[7]
mod_cor_AMYG_HIPPxAGE.c <- mod_ranef_corr$IDENT_SUBID[8]
mod_cor_AMYGxAGE.c_HIPPxAGE.c <- mod_ranef_corr$IDENT_SUBID[9]
mod_cor_HIPP.c_HIPPxAGE.c <- mod_ranef_corr$IDENT_SUBID[10]
```


